TimeVaryingDetection_mrmr

### 2024-11-13 code/R/TimeVaryingDetection_mrmr.Rmd
plot and visually compare model output

# Custom version of MLRG's mrmr package for mark-recapture data analysis

-   fit_model_custom() is a copy of mrmr's fit_model() function, changed only to call on a modified version of the underlying stan model.

@@

**Load packages**

```{r}
library(cmdstanr)
library(mrmr)
```

**Custom version of mrmr::fit_model()**

@@

# Custom fit function using the updated Stan model file
fit_model_custom <- function(
  data,
  seed = NULL,
  refresh = NULL,
  init = NULL,
  save_latent_dynamics = FALSE,
  output_dir = NULL,
  output_basename = NULL,
  sig_figs = NULL,
  chains = 4,
  parallel_chains = getOption("mc.cores", 1),
  chain_ids = seq_len(chains),
  threads_per_chain = NULL,
  opencl_ids = NULL,
  iter_warmup = 1000,
  iter_sampling = 1000,
  save_warmup = FALSE,
  thin = NULL,
  max_treedepth = NULL,
  adapt_engaged = TRUE,
  adapt_delta = NULL,
  step_size = NULL,
  metric = NULL,
  metric_file = NULL,
  inv_metric = NULL,
  init_buffer = NULL,
  term_buffer = NULL,
  window = NULL,
  fixed_param = FALSE,
  show_messages = TRUE,
  compile = TRUE,
  ...
) {
  # updated to use my custom stan model
  stan_file <- "JI_custom_twostate.stan" # specify local path

@@

@@

# set paths for Trapdoor captures and surveys
captures <- read.csv("data/54188_capture.csv")  
surveys <- read.csv("data/54188_survey.csv")

data_TVD_model <- mrmr::clean_data(captures, surveys)
```

# Fit the model using the custom function
fit_result <- fit_model_custom(data_TVD_model,
                                chains = 3, 
                                iter_sampling = 1000, 
                                iter_warmup = 400)

modTVD_summary <- fit_result$m_fit$summary()
View(modTVD_summary)

write.csv(modTVD_summary, "Trapdoor_custom_summary")

@@

**Now, run the classic mrmr model on the Trapdoor data to compare with my version**

# Fit the model using the custom function
Trapdoor1_mrmr <- mrmr::fit_model(data_TVD_model,
                                chains = 3, 
                                iter_sampling = 1000, 
                                iter_warmup = 400)

Trapdoor1_summary <- Trapdoor1_mrmr$m_fit$summary()
View(Trapdoor1_summary)

write.csv(Traodoor1_summary, "Trapdoor1_mrmr_summary.csv")

Trapdoor1_summary <- read.csv("Trapdoor1_mrmr_summary.csv")
View(trapdoor_summary)

@@

>>>>

# Test my custom stan model with MLRG's public data from site 54188

@@

**Load packages**
```{r}
library(cmdstanr)
library(mrmr)
library(ggplot2)
library(ggpubr)
```

## Modify mrmr::fit_model()

-   fit_model_custom() is a copy of mrmr::fit_model(), changed only to call on a modified version of the underlying stan model.

@@

# Custom fit function using the updated Stan model file
fit_model_custom <- function(
  data,
  seed = NULL,
  refresh = NULL,
  init = NULL,
  save_latent_dynamics = FALSE,
  output_dir = NULL,
  output_basename = NULL,
  sig_figs = NULL,
  chains = 4,
  parallel_chains = getOption("mc.cores", 1),
  chain_ids = seq_len(chains),
  threads_per_chain = NULL,
  opencl_ids = NULL,
  iter_warmup = 1000,
  iter_sampling = 1000,
  save_warmup = FALSE,
  thin = NULL,
  max_treedepth = NULL,
  adapt_engaged = TRUE,
  adapt_delta = NULL,
  step_size = NULL,
  metric = NULL,
  metric_file = NULL,
  inv_metric = NULL,
  init_buffer = NULL,
  term_buffer = NULL,
  window = NULL,
  fixed_param = FALSE,
  show_messages = TRUE,
  compile = TRUE,
  ...
) {
  # updated to use my custom stan model
  stan_file <- "/Users/johnimperato/Desktop/Mossy-Pond/code/stan/JI_custom_twostate.stan" # specify local path

@@

## Run 54188 data through my custom model

**Load and clean data**
```{r}
# set paths for 54188 captures and surveys
captures <- read.csv("/Users/johnimperato/Desktop/Mossy-Pond/data/54188_capture.csv")  
surveys <- read.csv("/Users/johnimperato/Desktop/Mossy-Pond/data/54188_survey.csv")

# run mrmr::clean_data()
data_54188 <- mrmr::clean_data(captures, surveys)
```

```{r}
# Fit the custom model using the modified function
custom_mod_54188 <- fit_model_custom(data_custom_mod_54188,
                                chains = 3, 
                                iter_sampling = 1000, 
                                iter_warmup = 400)

custom_mod_54188_summary <- custom_mod_54188$m_fit$summary()
View(custom_mod_54188_summary)

write.csv(custom_mod_54188_summary, "54188_custom_mod_summary")

**Plot abundance and recruitment**
```{r}

trapdoor_custom_abundance_plot <- mrmr::plot_model(custom_mod_54188, "abundance")
trapdoor_custom_recruitment_plot <- mrmr::plot_model(custom_mod_54188, "recruitment")

# arrange plots together
mrmr_custom_plots_54188 <- ggarrange(trapdoor_custom_abundance_plot, trapdoor_custom_recruitment_plot, ncol =1)

# add title
custom_annotated <- annotate_figure(mrmr_custom_plots_54188, top = text_grob("Time-Varying Detection mrmr", size = 14, face = "bold"))

## Run classic mrmr model on 54188 data

@@

# Fit the model using mrmr's built in function
classic_mod_54188 <- mrmr::fit_model(data_54188,
                                chains = 3, 
                                iter_sampling = 1000, 
                                iter_warmup = 400)

classic_mod_54188_summary <- classic_mod_54188$m_fit$summary()
View(classic_mod_54188_summary)

write.csv(classic_mod_54188_summary, "54188_classic_mod_summary.csv")

plot_model(classic_mod_54188, "abundance")
plot_model(classic_mod_54188, "recruitment")

@@

**Plot abundance and recruitment**
```{r}

trapdoor_classic_abundance_plot <- mrmr::plot_model(classic_mod_54188, "abundance")
trapdoor_classic_recruitment_plot <- mrmr::plot_model(classic_mod_54188, "recruitment")

mrmr_classic_plots_54188 <- ggarrange(trapdoor_classic_abundance_plot, trapdoor_classic_recruitment_plot, ncol =1)

# Add title
classic_annotated <- annotate_figure(mrmr_classic_plots_54188, top = text_grob("Classic mrmr", size = 14, face = "bold"))

@@

## Compare Model Output

**Visual Comparison**
```{r}

abund_recruit_comparison_plot <- ggarrange(classic_annotated, custom_annotated, ncol=2)
